 

### 2021-05-21 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Therefore, while under maximum $g_s$, leaves can maintain $T_{leaf}$ similar to $T_{air}$ at high solar-radiation loads until light saturation, during $g_s$ limitation solar radiation can drastically elevate $T_{leaf}$ above $T_{air}$ [@cavaleriColdbloodedForestsWarming2020].

High wind speed (Fig. `r fig_NEON_vertical`) decreases $T_{leaf}$ - $T_{air}$ due to large boundary layer conductance ($g_a$)  [@daudetWindSpeedLeaf1999a]. Therefore, under hot conditions, high wind speeds and high $g_s$ would enable cooling [Drake et al. 2018] but at lower wind conditions, thickness of air layer at the leaf surface curbs rapid leaf-to-air heat transfer, through smaller $g_a$, which increases $T_{leaf}$ - $T_{air}$ [@martin_boundary_1999; @leighInfluenceLeafSize2017; @robertsStomatalBoundaryLayerConductances1990a; @stokesBoundaryLayerConductance2006]

@@

**In addition, leaf size and morphology, which vary dramatically across the vertical gradient (reviewed in following section), impact $g_{a}$ and therefore leaf temperature.** Small leaves heat and cool faster due to higher $ga$ than larger leaves, therefore maintain lower $T_{leaf}$-$T_{air}$ than large leaves that experience greater $T_{leaf}$-$T_{air}$. With adequate water, species with smaller leaves and greater $gs$ will be most effective at cooling during high temperatures, whereas species with larger exposed-leaves can experience greater $T_{leaf}$-$T_{air}$ and heat stress [@songTreeSurfaceTemperature2020; @dongBiophysicalHomoeostasisLeaf2017; @bauerleSeparatingFoliarPhysiology2011; @leighInfluenceLeafSize2017; @songTreeSurfaceTemperature2020; @leuzingerTreeSpeciesDiversity2007]

>>>>

Therefore, while under maximum $g_s$, leaves can maintain $T_{leaf}$ similar to $T_{air}$ at high solar-radiation loads until light saturation, during $g_s$ limitation solar radiation can drastically elevate $T_{leaf}$ above $T_{air}$ [@fauset_differences_2018; @songTreeSurfaceTemperature2020].

High wind speed (Fig. `r fig_NEON_vertical`) decreases $T_{leaf}$ - $T_{air}$ due to large boundary layer conductance ($g_a$)  [@daudetWindSpeedLeaf1999a]. Therefore, under hot conditions, high wind speeds and high $g_s$ would enable cooling but at lower wind conditions, thickness of air layer at the leaf surface curbs rapid leaf-to-air heat transfer, through smaller $g_a$, which increases $T_{leaf}$ - $T_{air}$ [@martin_boundary_1999; @leighInfluenceLeafSize2017; @robertsStomatalBoundaryLayerConductances1990a; @stokesBoundaryLayerConductance2006]

@@

**In addition, leaf size and morphology, which vary dramatically across the vertical gradient (reviewed in following section), impact $g_{a}$ and therefore leaf temperature.** Small leaves heat and cool faster due to higher $ga$ than larger leaves, therefore maintain lower $T_{leaf}$-$T_{air}$ than large leaves that experience greater $T_{leaf}$-$T_{air}$. With adequate water, species with smaller leaves and greater $gs$ will be most effective at cooling during high temperatures, whereas species with larger exposed-leaves can experience greater $T_{leaf}$-$T_{air}$ and heat stress [@songTreeSurfaceTemperature2020; @dongBiophysicalHomoeostasisLeaf2017; @bauerleSeparatingFoliarPhysiology2011; @leighInfluenceLeafSize2017; @leuzingerTreeSpeciesDiversity2007]
